---
title: "Area_Height_Analysis"
author: "Steve Rolfe"
date: "5 November 2015"
output: html_document
---

Standard code to clear the system and install required libraries
```{r}
rm(list=ls())
library(ggplot2)
library(lubridate)
library(dplyr)
library(zoo)
```

Set up standard files names and the subset to plot
```{r}
setwd("C:/Users/Steve Rolfe/Google Drive/Phenomics/Area_Height")
input_file="./data/BR7_area_ht.csv"
plant_codes="./data/BR7_plant_codes.csv"

output_file="./output/BR7_area_ht_output.csv"


#subset of data to plot using grep format
#sset<-"BR7-0[1]1[1][1-4]"
sset<-"BR7-0[1-5]1[1-3][1-4]"
```

We use a rolling average to smooth out the noise - this sets the window size for this average
To calculate the maximum value we use a larger window
Finally we set the tolerance of how close a value has to be to this maximum to be the maximum
```{R}
#rolling window size - for the rolling average of water data
wsz<-3
#maximum window size - for the number of points from which to calculate the maximum
msz<-3
#how close to the maximum value does a value have to be to be counted as a maximum
csz<-0.95
```

Set up standard headers for the files - these need to be edited to be consistent
The BR codes are suppled by IBERS but we should include the ASSYT codes where possible
Standard headers should be
BR_code,date_time, measurements (with informative names)
Custom code will be needed to translate the BR_codes into useful information

Read in the data and check that the required headers are there
```{r}
#read in the data
htdata<-read.csv(input_file,header=TRUE,stringsAsFactors = FALSE)
if(is.object(htdata)==FALSE) {stop(c("Can't open file:",input_file))}

#check that there the required headings are there and throw an error if not
if(is.na(match("BR_code",colnames(htdata)))){stop("BR_code missing from input file header")}
if(is.na(match("date_time",colnames(htdata)))){stop("date_time missing from input file header")}
head(htdata)

#this file contains info about the plants used
plant_codes<-read.csv(plant_codes,header = TRUE,stringsAsFactors = FALSE)
if(is.object(plant_codes)==FALSE) {stop(c("Can't open file:",plant_codes))}
```

We  deal with any specific coding here about names of plants, spores etc.
Use the file plant_codes to keep the information organised
```{r}
#dpi offset from first data point
#it's good to express everything as dpi but 1st data point might not be dpi
#set the date of infection
dpi_datetime<-"2015-7-6 00:00:00"

#now set the factors manually
htdata$BR_code<-as.factor(htdata$BR_code)

#and set the plants,spores,etc
#the plant code is the sixth character in the BR7 string
htdata$plant<-as.factor(substr(htdata$BR_code,6,6))
#the spore concentration code is the 8th character
htdata$spore<-as.factor(substr(htdata$BR_code,8,8))
#the repeat is the 9th
htdata$rep<-as.factor(substr(htdata$BR_code,9,9))
head(htdata)

#now merge the plant names with the data frame
htdata<-merge(htdata,plant_codes,by.x="plant",by.y="plant_code")
head(htdata)

#now do some standard time processing
#use difftime to get the days since inoculation
htdata$dpi<-as.numeric(difftime(htdata$date_time,dpi_datetime,units="days"))
```

Always use a subset function to keep things under control for plotting

```{R}
subdata<-subset(htdata,grepl(sset,htdata$BR_code))
subdata<-droplevels(subdata)
#there's a lot of missing data (with NAs) so get rid
subdata<-na.omit(subdata)
```

Now plot the basic data
```{R}
head(subdata)
#plot the raw height data as a line
p1<-ggplot(data=subdata,aes(x=dpi,y=Plant_height_mm,color=spore,group=BR_code))+geom_line()
p1<-p1+scale_y_continuous(name="Height mm")
p1<-p1+facet_wrap(~plant_name)
p1

#plot the area
p2<-ggplot(data=subdata,aes(x=dpi,y=Area_mm2,color=spore,group=BR_code))+geom_line()
p2<-p2+scale_y_continuous(name="Area mm2")
p2<-p2+facet_wrap(~plant_name)
p2
```

Try a rolling average for the area data
```{R}
#sort the data so that the rolling means and cumulative values are correct
subdata<-subdata[with(subdata,order(BR_code,dpi)),]
#get the rolling average by group (window = wsz)
subdata$area_ravg<-ave(subdata$Area_mm2,subdata$BR_code,FUN=function(x) rollmean(x,k=wsz,fill = NA,na.pad=T))
head(subdata)

p2<-ggplot(data=subdata,aes(x=dpi,y=Area_mm2,color=spore,group=BR_code))+geom_point()
p2<-p2+geom_line(data=subdata,aes(x=dpi,y=area_ravg,colour=spore,group=BR_code))
p2<-p2+scale_y_continuous(name="Maximum Area mm2")
p2<-p2+facet_wrap(~plant_name)
p2
```

Now get some parameters from the data
```{R}
#get the maximum values  - sort by area rolling average
sort_subdata<-subdata[order(subdata$BR_code,subdata$area_ravg,na.last=TRUE,decreasing=TRUE),]
#create a vector to hold data by groups
grp<-group_by(sort_subdata,BR_code)
#get the mean of the msz th largest values
d<-summarise(grp,area_mean_ravg=mean(area_ravg[1:msz],na.rm=TRUE))
d
#merge
subdata<-merge(subdata,d,by="BR_code")


#plot the maximum area attained
p2<-ggplot(subdata,aes(x=spore,area_mean_ravg))
p2<-p2+geom_boxplot(aes(fill=spore),outlier.colour="red",outlier.size=3)
p2<-p2+scale_y_continuous(name="Maximum area (mm2)")
p2<-p2+facet_wrap(~plant_name)
p2
```

However, the maximum area is not the final area when measurements were taken
Get these as well
```{R}
#get the final values - sort on dpi
sort_subdata<-subdata[order(subdata$BR_code,subdata$dpi,na.last =TRUE,decreasing=TRUE),]
#create a vector to hold data by groups
grp<-group_by(sort_subdata,BR_code)
#get the mean of the msz th largest values
e<-summarise(grp,area_mean_final=mean(Area_mm2[1:msz],na.rm=TRUE))
e
subdata<-merge(subdata,e,by="BR_code")

#plot the maximum area attained
p2<-ggplot(subdata,aes(x=spore,area_mean_final))
p2<-p2+geom_boxplot(aes(fill=spore),outlier.colour="red",outlier.size=3)
p2<-p2+scale_y_continuous(name="Final area (mm2)")
p2<-p2+facet_wrap(~plant_name)
p2
```

Compare the two different measurements
```{R}
e<-merge(e,d,by="BR_code")
head(e)

#now set the factors manually - it's a bit ugly to set these values again rather than read them from subdata!
e$BR_code<-as.factor(e$BR_code)

#and set the plants,spores,etc
#the plant code is the sixth character in the BR7 string
e$plant<-as.factor(substr(e$BR_code,6,6))
#the spore concentration code is the 8th character
e$spore<-as.factor(substr(e$BR_code,8,8))
#the repeat is the 9th
e$rep<-as.factor(substr(e$BR_code,9,9))
head(e)

#now merge the plant names with the data frame
e<-merge(e,plant_codes,by.x="plant",by.y="plant_code")
head(e)

p3<-ggplot(e,aes(x=area_mean_final,y=area_mean_ravg))
p3<-p3+geom_point(aes(colour=spore,shape=plant_name,size=10))
p3<-p3+geom_smooth(method=lm)
p3

write.csv(e,file=output_file)
```


